\begin{abstract}
We address the problem of computing semantic differences between a program and a patched version of the program. Our goal is to obtain a precise characterization of the difference between program versions, or establish their equivalence when no difference exists.

We focus on computing semantic differences in numerical programs where the values of variables have no a-priori bounds, and use abstract interpretation to compute an over-approximation of program differences. Computing differences and establishing equivalence under abstraction requires abstracting relationships between variables in the original program and its patched version. Towards that end, we first construct a \emph{union program} in which these relationships can be identified, and then use a \emph{correlating abstract domain} to compute a sound approximation of these relationships. To establish equivalence between correlated variables and precisely capture differences, our domain has to represent non-convex information. To balance precision and cost of this representation, our domain may over-approximate numerical information as long as equivalence between correlated variables is preserved.

We have implemented our approach in a tool built on the LLVM compiler infrastructure and the APRON numerical abstract domain library, and applied it to a number of challenging real-world examples, including programs from the GNU core utilities, \TODO{Mozilla Firefox and the Linux Kernel}. We evaluate over 50 patches and show that for these programs, the tool often manages to establish equivalence, reports useful approximation of semantic differences when differences exists, and reports only a few false differences.
\end{abstract}
