\section{Related Work} \seclabel{Related}

Our work has been mainly inspired by recent work identifying program differencing as having vast security implications~\cite{BrumleyPoosankamSongZheng08,SongSunZhang09} as well as advancements made in the field of under-approximations of program equivalence~\cite{GodlinStrichman09, KawaguchiLahiriRebelo10, DwyerElbaumPerson08, EnglerRamos11}.

We rely on classic methods of abstract interpretation~\cite{CousotCousot77} for presenting an over approximating solution for semantic differencing and equivalence. To achieve this we devised a static analysis over a newly defined construct we call a correlating program. The idea of a correlating program is similar to that of
self-composition~\cite{BartheDArgenioRezk04, AikenTerauchi05} except that we compose two different programs in a interleaving designed to maintain a close correlation between them. The use of a correlating construct for differencing is novel as previous methods mainly use sequential composition~\cite{GodlinStrichman09, KawaguchiLahiriRebelo10, DwyerElbaumPerson08, EnglerRamos11}, disregarding possible program correlation.

We base our analysis on a relational abstraction ~\cite{CousotHalbwachs78, Mine07} that allows us to reason about variables of different programs. The abstraction is further refined towards a disjunctive domain, similar to trace partitioning~\cite{MauborgneRival07} and we use an equivalence based partitioning criteria, which is apt to our purposes.

Symbolic execution based methods~\cite{DwyerElbaumPerson08, EnglerRamos11} offer practical equivalence verification techniques for loop and recursion free programs with bound state space.



\TODO{ \cite{GS:DAC09} }


\paragraph{Determining corresponding components}

As suggested in \cite{Horwitz:PLDI90}, one possibility is to rely on the editing sequence that creates the new version from the original one. Another option is using various syntactic differencing algorithms as a base for computing correspondence tags.

\TODO{their idea for computing correspondence, is to minimize the ``size of change''. They have two different notions of size of change.}

\cite{ARRSY:CAV07} introduced a correlating heap semantics for verifying linearizability of concurrent programs. In their work, a correlating heap semantics is used to establish correspondence between a concurrent program and a sequential version of the program at specific linearization points.  