\section{Worklist}

\subsection{Points to Hammer}

\begin{enumerate}
\item a special kind of self composition, where correlated steps are kept together. This is particularly important when handling loops.
\end{enumerate}

\subsection{TODO}
\begin{enumerate}
\DONE{\item show me an example of where comparing abstract values at the end is not sound (e.g., we had some example with intervals that demonstrates this).}
\item run on uc-klee examples.
\item Consider describing each sub-state as a single (possibly looping) path of execution in both programs that originated from the same input.
\item in Overview, maybe show a bigger example that better shows how we benefit from partitioning
\item in Overview, talk about how merging early in paths may affect equivalence and differencing behaviors further along
\item Our evaluation includes applying each of our strategies along with each of the points. Intuitively, the results should range from least precise using the <Join-All,At-Join> strategy and point to most precise in the <Join-Equiv,At-Diff> scenario and this is indeed the case as we show in \secref{Evaluation} (not taking into account the no-partitioning scenario which is naturally most precise).
\item We also tested our tool on other works examples.
\end{enumerate}


\subsection{Questions}

\begin{enumerate}
\item how come you don't need the ``product program''?
\item what are the theorems that you provide? (no reason to have definitions if there are no theorems).
\item what abstract domains can we use as ``underlying domains'' for our abstraction? Do we have any particular requirements from the abstract domains (one requirement is being relational).
\item what makes a ``patched version of a program'' different from just saying ``a different program''? In other words - what are the requirements on the difference between $P$ and $P'$?
\item can we claim that our abstraction ``forgets'' paths along which equivalence is established, but keeps apart paths along which the is a difference, hoping that it will re-converge later?
\item (intuition only) what if we correlate badly and lose soundness? one can propose a 2 "trick" programs that correlating them our way gives a result that loses difference.
\item WHY DO WE CHECK DIFFERENCE ABOVE THE SUB-STATE LEVEL? doesn't that mean we compare different paths? isn't that bad?
\end{enumerate}

\subsection{Useful text fragments}

Differential static analysis is useful for regression debugging~\cite{TODO}, and may also lead to an automated approach for patch-based exploit generation~\cite{TODO}. Such automated exploit generation enables the organization releasing a patch to estimate the attack surface exposed by its release. Furthermore, it enables the organization releasing the patch to reduce vulnerability to manual and automated patch-based exploitation.


\begin{figure}
\begin{lstlisting}
if (input % 2 == 0) goto 2 else  goto 4;
s := input+2;
goto 5;
s := input+3;
if (s>input) goto 6 else goto ERROR;
ptr := realloc(ptr,s);
// use ptr[0], ptr[1], ... ptr[input-1]
\end{lstlisting}
\caption{Example from \cite{TODO}}
\end{figure} 